---
title: "Writing Assignment 3 - EDA"
author: "Melton Team"
date: "2/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data Cleaning

All of our data to this point has been hand typed and gathered in spreadsheets. We have been working to gather and fill in any gaps. One of the bright sides of this is that we don't have to undertake may cleaning steps as we can standardize the data as we enter it. We are limited by the sources we pull from. The most recent data is very complete, but looking 10 years ago we see much more spotty data as Microsoft's data reporting improved over time.We have been careful about how we enter the data to ensure no data entry errors. If we notice something that seems like a major outlier we check the input source and so far any outliers seem more to occur when the data was spottier or there was a change in reporting.

# Data Reshaping

As of this report all of our data is coming from the Carbon Disclosure Project(CDP). The only data reshaping that needs to be done is making sure to account for any changes in reporting over time. We are currently putting together a dataset on Data Centers but getting that data has been slow going. Once we have that completed data we have plans to correlate it with our existing CDP data in order to create our model.

# Key Metrics and Goals

We are trying to track carbon consumption over time. So metrics comparing carbon levels at various points in time and breaking down how that carbon is being emitted is at the heart of what we are doing. Other important metrics to keep in mind is how that carbon is being offset. How much and what type of offsets are being purchased can be correlated with our predictions for future carbon emissions leading to prediction in future cost of offsets. 


# Visualizations Over Time
## Global Emissions
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggpubr)
ccGlobal <- read.csv('Data/ClimateChangeGlobal.csv', header=FALSE)
ccGlobal = setNames(data.frame(t(ccGlobal[,-1])), ccGlobal[,1])
ccGlobal = ccGlobal[!is.na(ccGlobal[,1]),0:4]
ccGlobal = ccGlobal %>% rename(year = `Year (C6.1 - 6.5)`,
                               grossScope1 = `Gross global Scope 1 emissions (metric tons CO2e)`,
                               grossScope2Loc = `_Gross global Scope 2, location-based_ (metric tons CO2e)`,
                    grossScope2Mark = `Gross global Scope 2, market-based (metric tons CO2e)`) 

globalScope1 <- ggplot(ccGlobal, aes(x=year, y=grossScope1))+
  geom_point()+geom_line(linetype='dashed')+
  ggtitle("Gross global Scope 1 emissions (metric tons CO2e)")

globalScope2Loc <- ggplot(ccGlobal, aes(x=year, y=grossScope2Loc))+
  geom_point()+geom_line(linetype='dashed')+
  ggtitle("Gross global Scope 2, location-based (metric tons CO2e)")

globalScope2Mark <- ggplot(ccGlobal, aes(x=year, y=grossScope2Mark))+
  geom_point()+geom_line(linetype='dashed')+
  ggtitle("Gross global Scope 2, market-based (metric tons CO2e)")


ggarrange(globalScope1, globalScope2Loc, globalScope2Mark, nrow=2, ncol=2)
```

These graphs are at the heart of what we are doing. These are the graphs of total emissions over time, and this is what we are trying to predict. What we can see is that location based emissions are going up accross the board and Microsoft is opting to utilize buybacks to 'reduce' their emissions. 

## By Business Activity
```{r}
ccBusAct <- read.csv('Data/ClimateChangeBusinessActivity.csv')
ccBusAct = ccBusAct %>% rename(year = Year..C7.3.,
                    scope1 = Scope.1.emissions..metric.tons.CO2e.,
                    scope2Loc = Scope.2..location.based..metric.tons.CO2e.,
                    scope2Mark = Scope.2..market.based..metric.tons.CO2e.) %>%
  select(year,Activity,scope1,scope2Loc,scope2Mark) %>% 
  filter(!is.na(year),
         !is.na(Activity))

BAScope1Plot <- ggplot(ccBusAct, aes(x=year,y=scope1))+
  geom_point()+ facet_wrap(~Activity, scales='free')+
  ggtitle("Scope 1 Emissions")+
  geom_line(linetype='dashed')

BAScope2LocPlot <- ggplot(ccBusAct, aes(x=year,y=scope2Loc))+
  geom_point()+ facet_wrap(~Activity, scales='free')+
  ggtitle("Scope 2 (Location-Based) Emissions")+
  geom_line(linetype='dashed')

BAScope2MarkPlot <- ggplot(ccBusAct, aes(x=year,y=scope2Mark))+
  geom_point()+ facet_wrap(~Activity, scales='free')+
  ggtitle("Scope 2 (Market-Based) Emissions")+
  geom_line(linetype='dashed')

BAScope1Plot
BAScope2LocPlot
BAScope2MarkPlot
```
This data is in large part why we decided to focus on datacenters. As we can see from these graphs. Emissions are going down in almost all areas over time, but data center emissions are growing exponentially. So tracking data center emissions in large part is well representative of all emissions.

## By Greenhouse Gas
```{r}
ccGas <- read.csv('Data/CCGas.csv')
ccGas = ccGas %>% rename(year = Year..C7.1a.,
                         emissions = Scope.1.Emissions..metric.tons.CO2e.)

GasPlot <- ggplot(ccGas, aes(x=year, y=emissions))+
  geom_point()+ facet_wrap(~Gas, scales='free')+
  ggtitle("Emissions by Greenhouse Gas")+
  geom_line(linetype='dashed')
GasPlot
```




## By Area
```{r}
ccArea <- read.csv('Data/CCArea.csv')
ccArea = ccArea %>% rename(year = Year..C7.2.7.5.,
                         scope1 = Scope.1.Emissions..metric.tons.CO2e.,
                         scope2Loc = Scope.2..location.based..metric.tons.CO2e.,
                         scope2Mark = Scope.2..market.based..metric.tons.CO2e.,
                         ElHeStCool = Purchased.and.consumed.electricity..heat..steam.or.cooling..MWh.,
                         lowCarbElHeStCool = Purchased.and.consumed.low.carbon.electricity..heat..steam.or.cooling.accounted.for.in.Scope.2.market.based.approach..MWh.)%>% 
  filter(!is.na(Area))

ccArea$Area[ccArea$Area=="Latin America (LATAM)"]="Latin America and Caribbean (LAC)"
ccArea = ccArea[!(ccArea$Area=="Canada"|ccArea$Area=="United States of America"),]
AreaScope1Plot <- ggplot(ccArea, aes(x=year, y=scope1))+
  geom_point()+ facet_wrap(~Area, scales='free')+
  ggtitle("Scope 1 Emissions by Location")+
  geom_line(linetype='dashed')
AreaScope1Plot
```
```{r}
AreaScope2LocPlot <- ggplot(ccArea, aes(x=year, y=scope2Loc))+
  geom_point()+ facet_wrap(~Area, scales='free')+
  ggtitle("Scope 2 (location-based) Emissions by Location")+
  geom_line(linetype='dashed')
AreaScope2LocPlot
```
This graph is showing actual emissions within scope2 based on their location. We see a steady increase across the bored in all major regions. One thing to note on these graphs is scale. The scale is not the same between these graphs and while representative in the change in their area, the fluctions you see in Latin America is actually fairly small. 

```{r}
AreaScope2MarkPlot <- ggplot(ccArea, aes(x=year, y=scope2Mark))+
  geom_point()+ facet_wrap(~Area, scales='free')+
  ggtitle("Scope 2 (market-based) Emissions by Location")+
  geom_line(linetype='dashed')
AreaScope2MarkPlot
```
These graphs don't paint a complete picture of total emissions emitted by area. These are graphs of market based emissions, so these are showing emissions after buybacks, which largely started in 2016. What it does clearly show is how most buybacks are currently taking place in Europe

```{r}
ggplot(ccArea, aes(x=year))+
  geom_point(aes(y=ElHeStCool), color="blue")+
  geom_line(aes(y=ElHeStCool), color="blue",linetype="dashed")+
  geom_point(aes(y=lowCarbElHeStCool), color="purple")+
  geom_line(aes(y=lowCarbElHeStCool), color="purple",linetype="dashed")+
  facet_wrap(~Area, scales='free')+
  ggtitle("Purchased and Consumed Electricity, Heat, Steam, or Cooling by Location")
```

## By Consumption Type
```{r}
ccConType <- read.csv('Data/EnergyConsumptionOverview.csv')
ccConType = ccConType %>% rename(year = Year..C8.2.,
                         conType = Consumption.Type..purchased.or.aquired.,
                         MWhRenew = MWh.from.renewable.sources,
                         MWhNonRenew = MWh.from.non.renewable.sources,
                         MWhTotal = Total..renewable.and.non.renewable..MWh) %>%
  select(year,conType,MWhRenew,MWhNonRenew,MWhTotal) %>%
  filter(!is.na(conType))

ggplot(ccConType, aes(x=year))+
  geom_point(aes(y=MWhTotal), color="black")+
  geom_line(aes(y=MWhTotal), color="black",linetype="dashed")+
  geom_point(aes(y=MWhRenew), color="blue")+
  geom_line(aes(y=MWhRenew), color="blue",linetype="dashed")+
  geom_point(aes(y=MWhNonRenew), color="purple")+
  geom_line(aes(y=MWhNonRenew), color="purple",linetype="dashed")+
  facet_wrap(~conType, scales='free')+
  ggtitle("")
```
In this graph we are looking at the breakdown of uses of energy consuming carbon. Black is the total, Blue is renewables and purple is non renewables. We see a start in the color change in 2018 because they changed how they were reporting MWh Totals that year. What we can see is that electricity and fuel are the largest sources. We see an effort by Microsoft to make all of their electricity come from renewable sources at the same time as they made a net zero pledge.


## By Fuel Breakdown

```{r}
FuelBreakdownType <- read.csv('Data/FuelUseBreakdown.csv')
FuelBreakdownType = FuelBreakdownType %>% rename(year = Year..C8.2c.,
                         fuelType = Fuel.Type,
                         totalFuel = Total.fuel.MWh.consumed.by.the.organization,
                         MWhElectric = MWh.fuel.consumed.for.self.generation.of.electricity,
                         MWhHeat = MWh.fuel.consumed.for.self.generation.of.heat,
                         EmissFactor = Emission.factor..metric.tons.CO2e.per.MWh.) %>%
  select(year,fuelType,totalFuel,MWhElectric,MWhHeat,EmissFactor) %>%
  filter(!is.na(fuelType))

 

ggplot(FuelBreakdownType, aes(x=year))+
  geom_point(aes(y=totalFuel), color="black")+
  geom_line(aes(y=totalFuel), color="black",linetype="dashed")+
  geom_point(aes(y=MWhElectric), color="blue")+
  geom_line(aes(y=MWhElectric), color="blue",linetype="dashed")+
  geom_point(aes(y=MWhHeat), color="purple")+
  geom_line(aes(y=MWhHeat), color="purple",linetype="dashed")+
  geom_point(aes(y=EmissFactor), color="red")+
  geom_line(aes(y=EmissFactor), color="red",linetype="dashed")+
  facet_wrap(~fuelType, scales='free')+
  ggtitle("Fuel Type Breakdown")
```
This graph helps to break down where the non-renewable are going between Heat and electricity, blue and purple respectively with black being the total. Most of which is going towards electricity. aside from Fuel Oil Number 2 and Natural Gas. We can also see an effort by Microsoft to reduce their usage over time in most categories. 


## By Water Discharge

```{r}
WaterDischarge <- read.csv('Data/WaterDischarge.csv')
WaterDischarge = WaterDischarge %>% rename(year = Year..W1.2i.,
                         waterType = Type,
                         Volume = Volume..megaliters.year.,) %>%
  select(year,waterType,Volume) %>%
  filter(!is.na(waterType))

 

ggplot(WaterDischarge, aes(x=year))+
  geom_point(aes(y=Volume), color="blue")+
  geom_line(aes(y=Volume), color="blue",linetype="dashed")+
  facet_wrap(~waterType, scales='free')+
  ggtitle("Water Discharge by Type")
```
Water discharged by Microsft can effect local habitats and have an indirect effect on climate change. From this data it appears that discharge has been mostly stable or near 0. I don't see us using this as a factor for future carbon emissions.

## By Water Withdrawal
```{r}
WaterWithdrawal <- read.csv('Data/WaterWithdrawal.csv')
WaterWithdrawal = WaterWithdrawal %>% rename(year = Year..W1.2h.,
                         waterType = Water.Source,
                         Volume = Volume..megaliters.year.,) %>%
  select(year,waterType,Volume) %>%
  filter(!is.na(waterType))

 

ggplot(WaterWithdrawal, aes(x=year))+
  geom_point(aes(y=Volume), color="blue")+
  geom_line(aes(y=Volume), color="blue",linetype="dashed")+
  facet_wrap(~waterType, scales='free')+
  ggtitle("Water Withdrawal by Type")
```



# Conclusions So Far
From the data we have it is clear that electricity usage and carbon emmisons by microsoft is on the rise. With Microsoft creating 50-100 new data centers every year wether though building, buying or leasing, it is no wonder that emissions in those areas are rising the fastest. By utilzing predictions surrounding datacenters we can give a good estimate of how carbon consumption is likely to rise by Microsoft and how that is going to translate to future buybacks. 

